`methyvim`: Targeted, Robust, and Model-free Analysis of Differential

### 2018-06-29 paper/paper_BiocF1000.Rmd
rework paper; update methods

---
title: '`methyvim`: Targeted and Model-free Differential Methylation Analysis in
  R'
author:
  - name: Nima S. Hejazi
    affiliation:
      - Group in Biostatistics, University of California, Berkeley
      - Center for Computational Biology, University of California, Berkeley
    orcid: 0000-0002-7127-2789
    email: nhejazi@berkeley.edu
  - name: Rachael V. Phillips
    affiliation:
      - Group in Biostatistics, University of California, Berkeley
    orcid: 0000-0002-8474-591X
  - name: Alan E. Hubbard
    affiliation:
      - Group in Biostatistics, University of California, Berkeley
    orcid: 0000-0002-3769-0127
  - name: Mark J. van der Laan
    affiliation:
      - Group in Biostatistics, University of California, Berkeley
      - Department of Statistics, University of California, Berkeley
    orcid: 0000-0003-1432-5511
abstract: |
  We present `methyvim`, an R package implementing a general algorithm for the
  nonparametric estimation of treatment effects on DNA methylation at CpG sites
  throughout the genome, complete with straightforward statistical inference for
  such estimates. The approach leverages variable importance measures derived
  from statistical parameters arising causal inference, defined in such a manner
  that they may be used to obtain targeted estimates of the relative importance
  of individual CpG sites with respect to a binary treatment assigned at the
  phenotype level. The procedure implemented is computationally efficient,
  incorporating a preliminary screening step to isolate a subset of sites for
  which there is cursory evidence of differential methylation as well as a
  unique multiple testing correction to control the False Discovery Rate with
  the same rigor as if all sites were tested. This technique for analysis of
  differentially methylated positions provides an avenue to incorporate flexible
  state-of-the-art machine learning algorithms into the estimation of
  differential methylation effects without the loss of interpretable statistical
  inference.
keywords: DNA methylation, differential methylation, epigenetics, causal
  inference, statistical variable importance, machine learning, targeted minimum
  loss-based estimation
bibliography: references.bib
output: BiocWorkflowTools::f1000_article
---

```{r setup, echo=FALSE}
options(scipen=999)
```

@@

In this paper, we briefly present an alternative to such statistical analysis
approaches in the form of a nonparametric estimation procedure that provides
simple and readily interpretable statistical inference, discussing at length a
recent implementation of the methodology in the `methyvim` R package. Inspired
by recent advances in statistical causal inference and machine learning, we
provide a computationally efficient technique for obtaining targeted estimates
of nonparametric _variable importance measures_ (VIMs) [@vdl2006statistical],
estimated at a set of pre-screened CpG sites, controlling for the False
Discovery Rate (FDR) as if all sites were tested. Under standard assumptions
(e.g., identifiability, strong ignorability) [@pearl2009causality], targeted
minimum loss-based estimators of regular asymptotically linear estimators have
sampling distributions that are asymptotically normal, allowing for reliable
point estimation and the construction of Wald-style confidence intervals
[@vdl2011targeted, @vdl2018targeted]. As counterfactuals are contrary-to-fact --
that is, defined in terms of missing data -- in the context of DNA methylation,
we define the counterfactual outcomes under a binary treatment as the observed
methylation (whether Beta- or M-) values a CpG site would have if all subjects
were given the treatment and the methylation values a CpG site would have if
treatment were withheld from all subjects. Although these counterfactual
outcomes are, of course, impossible to observe, they have statistical analogs
that may be reliably estimated from observed data [@pearl2009causality]. We
describe an algorithm that incorporates the estimation of VIMs using _targeted
minimum loss-based estimation_ (__TMLE__) [@vdl2006targeted], deferring detailed
description and analysis of the statistical methodology to work outside the
scope of the present manuscript. This methodology assesses the individual
importance of a given CpG site by utilizing state-of-the-art machine learning
algorithms in the estimation of a targeted VIM, initially proposed and explored
in [@tuglus2008targeted]. In the present work, we focus on a software package,
`methyvim`, that implements a variant of this methodology, specifically tailored
to differential methylation analysis.

For a general discussion of the framework of targeted minimum loss-based
estimation and the role this approach plays in statistical causal inference, the
interested reader is invited to consult @vdl2011targeted and @vdl2018targeted.
For a more general introduction to (statistical) causal inference,
@pearl2009causality and @hernan2018causal may be of interest.

@@

We consider an observed data structure $O = (W_k, A, Y_j)$, where $Y_j, j = 1,
\ldots J$ represents a given CpG site of interest, $A \in \{0, 1\}$ represents a
binary phenotype-level treatment, and $W_k, k = 1, \ldots K(j)$ is a matrix of
the observed values of CpG sites in the same neighborhood as $j$ as well as any
potential phenotype-level confounders (e.g., sex, age). We consider having
access to measurements on a large number $G$ of CpG sites (e.g., $850,000$, as
measured by the Illumina MethylationEPIC BeadChip arrays), and as a matter of
practicality, let $J \leq G$, so that $J$ indexes only those CpG sites that pass
an arbitrary pre-screening procedure; note that equality is only obtained in the
case that all sites pass the pre-screening procedure. Further, we let $K(j)$
index the set of all CpG sites so that $k$, the realization of $K(j)$ for a
given CpG site $j$, is a set of the CpG sites that neighbor the CpG site $j$,
where the definition of a neighborhood is left as user-specified; thus, $W_k$ is
a matrix of observed methylation values for the set of CpG sites $k$ that
neighbor a particular target site $j$ as well as any potential phenotype-level
confounders, where the latter does not vary between realizations of $K(j)$. As
a variable importance measure, we see to estimate a target parameter ($\psi_j$)
that quantifies the effect of changing a treatment $A$ on the methylation of a
CpG site $j$, accounting for any potential confounding from phenotype-level
covariates and the observed methylation at the set of sites $k$ that neighbor
the site $j$. After a complete run of the procedure, we have access to $j$ VIM
estimates $\psi_j$, each corresponding to the degree to which a given CpG site
$j$ was responsive to a change in the binary treatment $A$, after controlling
for potential confounding. When machine learning estimators are employed in
estimating $\psi_j$, this produces a nonparametric variable importance measure
of differential methylation, allowing for the identification of differentially
methylated positions (DMPs). We note that $\psi(j)$, for any given $j$, is a
function of the _true_ data-generating distribution; so, to formalize, we
consider observing $n$ iid copies of $O$ (i.e., $O_1, \ldots, O_n$), where we
have $O \sim P \in \mathcal{M}$, which is to say that the random variable $O$ is
governed by an unknown probability distribution $P$, assumed only to reside in a
nonparametric statistical model $\mathcal{M}$ that places no restrictions on the
data-generating process.

@@

The core functionality of this package is made available via the eponymous
`methyvim` function, which implements a statistical algorithm designed to
compute targeted estimates of VIMs, defined in such a way that the VIMs
represent parameters of scientific interest in computational biology
experiments; moreover, these VIMs are defined such that they may be estimated in
a manner that is very nearly assumption-free, that is, within a _fully
nonparametric statistical model_. __The statistical algorithm consists in
several major steps__:

1. Pre-screening of genomic sites is used to isolate a subset of sites for
   which there is cursory evidence of differential methylation. For the sake of
   computational feasibility, targeted minimum loss-based estimates of VIMs are
   computed only for this subset of sites. Currently, the available screening
   approach adapts core routines from the
   [`limma`](http://bioconductor.org/packages/limma) R
   package. Future releases will support functionality from other packages
   (e.g., [`randomForest`](https://CRAN.R-project.org/package=randomForest),
   [`tmle.npvi`](https://CRAN.R-project.org/package=tmle.npvi)). Following the
   style of the function for performing screening via `limma`, users may write
   their own screening functions and are invited to contribute such functions to
   the core software package by opening pull requests at the GitHub repository.

2. Nonparametric estimates of VIMs, for the specified target parameter, are
   computed at each of the CpG sites passing the screening step. The VIMs are
   defined in such a way that the estimated effects is of an exposure/treatment
   on the methylation status of a target CpG site, controlling for the observed
   methylation status of the neighbors of that site. Currently, routines are
   adapted from the [`tmle`](https://CRAN.R-project.org/package=tmle) R package.
   Future releases will support doubly-robust estimates of these VIMs (via the
   [`drtmle`](https://cran.r-project.org/package=drtmle) package)
   and add parameters for continuous treatments/exposures (via the
   [`tmle.npvi`](https://CRAN.R-project.org/package=tmle.npvi) and
   [`txshift`](https://github.com/nhejazi/txshift) R packages).

3. Since pre-screening is performed prior to estimating VIMs, we make use of a
   multiple testing correction uniquely suited to such settings. Due to the
   multiple testing nature of the estimation problem, a variant of the Benjamini
   and Hochberg procedure for controlling the False Discovery Rate (FDR) is
   applied [@benjamini1995controlling]. Specifically, we apply a modified
   marginal Benjamini and Hochberg step-up False Discovery Rate controlling
   procedure for multi-stage analyses (FDR-MSA), which has established
   theoretical guarantees to control the FDR at the same rate as if all sites
   were tested (i.e., without screening) [@tuglus2009modified].

@@

Support for continuous-valued treatments or exposures is _planned but not yet
available_. Future releases will allow users to assess continuous-valued
treatments by relying on parameters estimable through implementations available
in the following software packages:

@@

We refer the reader to the package vignette, "`methyvim`: Targeted Data-Adaptive
Estimation and Inference for Differential Methylation Analysis," included in any
distribution of the software package, for further details.

@@

```{r setup-minfidata}
suppressMessages(library(minfiData))
data(MsetEx)
mset <- mapToGenome(MsetEx)
grs <- ratioConvert(mset)
grs
```
We can create an object of class `methytmle` from any `GenomicRatioSet` object
simply invoking the S4 class constructor `.methytmle`:

@@

```{r fdr-msa}
fdr_p <- fdr_msa(pvals = methyvim_cancer_ate@vim$pval,
                 total_obs = nrow(methyvim_cancer_ate))
```

@@

```{r methyvim-pvals-raw}
plot(methyvim_cancer_ate, type = "raw_pvals")
```

```{r methyvim-pvals-fdr}
plot(methyvim_cancer_ate, type = "fdr_pvals")
```

@@

```{r methyvim-volcano}
methyvolc(methyvim_cancer_ate)
```

@@

```{r methyvim-heatmap}
methyheat(methyvim_cancer_ate, smooth.heat = TRUE, left.label = "none")
```

@@

No competing interests were disclosed.

>>>>

---
title: '`methyvim`: Targeted, Robust, and Model-free Analysis of Differential
  Methylation in R'
author:
  - name: Nima S. Hejazi
    affiliation:
      - Group in Biostatistics, University of California, Berkeley
      - Center for Computational Biology, University of California, Berkeley
    orcid: 0000-0002-7127-2789
    email: nhejazi@berkeley.edu
  - name: Rachael V. Phillips
    affiliation:
      - Group in Biostatistics, University of California, Berkeley
    orcid: 0000-0002-8474-591X
  - name: Alan E. Hubbard
    affiliation:
      - Group in Biostatistics, University of California, Berkeley
    orcid: 0000-0002-3769-0127
  - name: Mark J. van der Laan
    affiliation:
      - Group in Biostatistics, University of California, Berkeley
      - Department of Statistics, University of California, Berkeley
    orcid: 0000-0003-1432-5511
abstract: |
  We present `methyvim`, an R package implementing a general algorithm for the
  nonparametric estimation of treatment effects on DNA methylation at CpG sites
  throughout the genome, complete with straightforward statistical inference for
  such estimates. The approach leverages variable importance measures derived
  from statistical parameters arising in causal inference, defined in such a
  manner that they may be used to obtain targeted estimates of the relative
  importance of individual CpG sites with respect to a binary treatment assigned
  at the phenotype level. The procedure implemented is computationally
  efficient, incorporating a preliminary screening step to isolate a subset of
  sites for which there is cursory evidence of differential methylation as well
  as a unique multiple testing correction to control the False Discovery Rate
  with the same rigor as if all sites were tested. This technique for analysis
  of differentially methylated positions provides an avenue to incorporate
  flexible state-of-the-art machine learning algorithms into the estimation of
  differential methylation effects without the loss of interpretable statistical
  inference.
keywords: DNA methylation, differential methylation, epigenetics, causal
  inference, statistical variable importance, machine learning, targeted minimum
  loss-based estimation
bibliography: references.bib
output: BiocWorkflowTools::f1000_article
---

@@

In this paper, we briefly present an alternative to such statistical analysis
approaches in the form of a nonparametric estimation procedure that provides
simple and readily interpretable statistical inference, discussing at length a
recent implementation of the methodology in the `methyvim` R package. Inspired
by recent advances in statistical causal inference and machine learning, we
provide a computationally efficient technique for obtaining targeted estimates
of nonparametric _variable importance measures_ (VIMs) [@vdl2006statistical],
estimated at a set of pre-screened CpG sites, controlling for the False
Discovery Rate (FDR) as if all sites were tested. Under standard assumptions
(e.g., identifiability, strong ignorability) [@pearl2009causality], targeted
minimum loss-based estimators of regular asymptotically linear estimators have
sampling distributions that are asymptotically normal, allowing for reliable
point estimation and the construction of Wald-style confidence intervals
[@vdl2011targeted, @vdl2018targeted]. In the context of DNA methylation studies,
we define the counterfactual outcomes under a binary treatment as the observed
methylation (whether Beta- or M-) values a CpG site would have if all subjects
were administered the treatment and the methylation values a CpG site would have
if treatment were withheld from all subjects. Although these counterfactual
outcomes are, of course, impossible to observe, they do have statistical analogs
that may be reliably estimated (i.e., identified) from observed data under a
small number of untestable assumptions [@pearl2009causality]. We describe an
algorithm that incorporates, in its final step, the use _targeted minimum
loss-based estimators_ (__TMLE__) [@vdl2006targeted] of a given VIM of interest,
though we defer rigorous and detailed descriptions of this aspect of the
statistical methodology to work outside the scope of the present manuscript
[@vdl2006targeted, @vdl2011targeted, @vdl2018targeted]. The proposed methodology
assesses the individual importance of a given CpG site, as a proposed measure of
differential methylation, by utilizing state-of-the-art machine learning
algorithms in deriving targeted estimates and robust inference of a VIM, as
considered more broadly for biomarkers in @bembom2009biomarker and
@tuglus2011targeted. In the present work, we focus on the `methyvim` software
package, available through the Bioconductor project [@gentleman2004bioconductor,
@huber2015orchestrating] for the R language and environment for statistical
computing [@R], which implements a particular realization of this methodology
specifically tailored for the analysis and identification of differentially
methylated positions (DMPs).

For an extended discussion of the general framework of targeted minimum
loss-based estimation and detailed accounts of how this approach may be brought
to bear in developing answers to complex scientific problems through tatistical
and causal inference, the interested reader is invited to consult
@vdl2011targeted and @vdl2018targeted. For a more general introduction to causal
inference, @pearl2009causality and @hernan2018causal may be of interest.

@@

We consider an observed data structure, on a single experimental subject (e.g.,
a patient), $O = (W, A, (Y(j) : j))$, where $(Y(j) : j = 1, \ldots J)$ is the
set of CpG sites measured by the assay in question, $A \in \{0, 1\}$ represents
a binary phenotype-level treatment, and $W$ is a vector of the phenotype-level
baseline covariates that are potential confounders (e.g., age, sex). We consider
having access to measurements on a large number $J$ of CpG sites (e.g.,
$850,000$, as measured by the Illumina _MethylationEPIC_ BeadChip arrays).
Further, let $K: j \to S_j$ be a procedure that assigns to a given CpG site $j$
a set of neighbors $S_j$ -- that is, $S_j$ is a collection of indices of the
neighbors of $j$ just as $j$ indexes the full set of CpG sites; moreover, since
$Y(j)$ is the measured methylation value for a given CpG site $j$, $Y(S_j)$ is
the measured methylation values at the neighbors $S_j$ of $j$. Note that the
definition of a neighborhood $\{j, S_j\}$ is left vague so as to facilitate the
use of user-specified strategies in implementation. We consider the case of
observing $n$ iid copies of $O$ (i.e., $O_1, \ldots, O_n$), where $O \sim P \in
\mathcal{M}$, which is to say that the random variable $O$ is governed by an
unknown probability distribution $P$, assumed only to reside in a nonparametric
statistical model $\mathcal{M}$ that places no restrictions on the
data-generating process.

## Variable Importance Measure

With the data structure above in hand, we let the estimand of interest be a
$j$-specific variable importance measure (VIM) $\Psi_j(P)$, which is defined
through the _true_ (and unknown) probability distribution $P$. As a motivating
example, consider the case where $\Psi_j(P)$ is
\begin{equation}\label{vim_param}
  \Psi_j(P) = \mathbb{E}_P(\mathbb{E}_P(Y(j) \mid A = 1, W, Y(S_j)) -
    \mathbb{E}_P \mathbb{E}_P(Y(j) \mid A = 0, W, Y(S_j)),
\end{equation}
where $Y(S_j)$ is the subvector of $Y$ that contains the measured methylation
values of the _neighboring sites_ of $j$ (but not site $j$ itself). This
parameter of interest is a variant of the _average treatment effect_, which has
been the subject of much attention in statistical causal inference **CITATION
NEEDED**. As a measure of variable importance, we seek to estimate this target
parameter ($\Psi_j(P)$), which quantifies the effect of changing a treatment $A$
on the methylation $Y(j)$ of a CpG site $j$, accounting for any potential
confounding from phenotype-level covariates $W$ and observed methylation
$Y(S_j)$ at the neighboring sites $S_j$ of $j$. We propose a procedure that
estimates this target parameter $\Psi_j(P)$ across all CpG sites of interest
$j: 1, \ldots, J$. When data-adaptive regression procedures (i.e., machine
learning) are employed in estimating $\Psi_j(P)$, this produces a nonparametric
variable importance measure of differential methylation, allowing for the
identification of differentially methylated positions (DMPs) while avoiding many
assumptions common in the use of standard parametric regression procedures. We
propose estimating $\Psi_j(P)$ via targeted minimum loss-based estimation, which
allows for data-adaptive regression procedures to be employed in a
straightforward manner.

## Pre-Screening Procedure

As a matter of practicality, we consider only estimating the target VIM
$\Psi_j(P)$ at a subset of CpG sites, so that $j: 1, \ldots J$ does not, in fact
span the full set of assayed CpG sites. In order to determine this subset of CpG
sites, we propose the use of a pre-screening procedure, which need be nothing
more than a method for differential methylation analysis that is computationally
less demanding than the method proposed here. Formally, let the pre-screening
procedure $\theta(j): j \to \{0, 1\}$, which is to say that $\theta$ takes as
input a CpG site $j$ and returns a binary decision rule of whether to include
CpG site $j$ in a subset to be evaluated further or not. Details on the
implementation of pre-screening approaches is available in the sequel.

## Reducing Neighbors via Clustering

Due to the data-adaptive nature of the regression procedures employed in
evaluating the target VIM $\Psi_j(P)$ via targeted minimum loss-based
estimation, it is possible the a given CpG site $j$ may have _too many_
neighbors $S_j$ to be controlled for in the estimation procedure. Heuristically,
the inclusion of too many neighbors when controlling for potential confounders
may lead to instability in the estimates produced. In such cases, we propose and
implement the use of a clustering technique (e.g., Partitioning Around Medoids)
to select a _representative_ subset of neighbors. Formally, there is likely no
best choice of a specific clustering algorithm, so we leave this aspect of the
proposed algorithm as flexible for the user [@kleinberg2003impossibility].

## Targeted Minimum Loss-Based Estimation and Inference

Given the choice of target parameter $\Psi_j(P)$, we propose the use of targeted
minimum loss-based estimation (TMLE) to construct and evaluate an estimator
$\psi_{n, j}$ of $\Psi_j(P)$. In the case of our motivating example, where the
target VIM is the average treatment effect, we make use of a TML estimator of
this parameter, which has been implemented for a general case in the
[`tmle`](https://CRAN.R-project.org/package=tmle) R package [@gruber2011tmle].
Generally speaking, a TML estimator is constructed from a few simple components:
(1) an estimator of the propensity score [@rosenbaum1983central] (often denoted
by $g(A, W)$), (2) an estimator of the outcome regression (often denoted $Q(A,
W)$), and (3) a targeting step applied to the initial estimator constructed from
the aforementioned components [@gruber2009targeted]. In the sequel, we describe
a few of the key mathematical properties of TML estimators relevant for
statistical inference, deferring extended discussion of the technical details to
@vdl2011targeted and @vdl2018targeted; moreover, as a matter of practicality, we
make use of the [`tmle`](https://CRAN.R-project.org/package=tmle)  R package of
@gruber2011tmle as a backend for computing estimators of relevant target
parameters in our software package, so the interested user is encouraged to
discuss the documentation of that software package for further details.

As noted previously, data-adaptive regression techniques may be used in the
construction of TML estimators. This is done by generating a stacked regression
model from a given set of machine learning algorithms via the Super Learner
algorithm [@wolpert1992stacked, breiman1996stacked, vdl2007super], which ensures
that the resulting ensemble learner satisifies important optimality properties
with respect to cross-validation [@vdl2004asymptotic, @dudoit2005asymptotics,
@vdl2003unified]. Importantly, TML estimators are well suited for inference,
having an asymptotically normal limiting distribution
[@tsiatis2007semiparametric, @vdl2006targeted] $\sqrt{n}(\psi_{n, j} -
\Psi_j(P)) \sim N(0, \sigma_{\text{EIF}}^2)$, where $\sigma_{\text{EIF}}^2$ is
the variance of the efficient influence function of the target parameter
evaluated at the observed data. This allows for confidence intervals to be
constructed and and hypothesis tests (i.e., $H_{0, j}: \Psi_j(P) = 0$) to be
performed in a straightforward manner.

## Correction for Multiple Testing

Given that we seek to estimate $\Psi_j(P)$ for a possibly large number of CpG
sites $(j: 1, \ldots J)$, the need to perform corrections for multiple testing
is clear. In order to curb the potential for false discoveries, we recommend the
use of the Benjamini and Hochberg procedure for controlling the False Discovery
Rate (FDR) [@benjamini1995controlling]; however, as the proposed procedure
involves a pre-screening step, naive application of the Benjamini and Hochberg
procedure (BH) is invalid -- instead, we rely on a modification of the procedure
to control the FDR, with established theoretical guarantees when pre-screening
is employed [@tuglus2009modified]. In brief, the modified marginal Benjamini and
Hochberg procedure to control the FDR under pre-screening works by applying the
standard BH procedure to a padded vector of p-values -- that is, letting $J$ be
the number of CpG sites tested and $K$ the number of CpG sites filtered out (so
that $J + K = P$, where $P$ is the original dimension of the genomic assay), the
modified marginal BH procedure is the application of the original BH procedure
to a vector of p-values, composed of the $J$ p-values from performing $J$
hypothesis tests (of the form $H_{0, j}: \Psi_j(P) = 0$) and $K$ additional
p-values automatically set to $1$ (for the hypothesis tests not performed on
account of pre-screening). This procedure is guaranteed to control the FDR at
the same desired rate when pre-screening is performed whereas naive application
of the BH procedure fails to do so.

@@

The core functionality of this package is made available via the eponymous
`methyvim` function, which implements a statistical algorithm designed to
compute targeted estimates of VIMs, defined in such a way that the VIMs
represent parameters of scientific interest in computational biology
experiments; moreover, these VIMs are defined such that they may be estimated in
a manner that is very nearly assumption-free, that is, within a _fully
nonparametric statistical model_. __The statistical algorithm consists in the
several major steps elaborated upon above and summarized below__:

1. _Pre-screening_ of genomic sites is used to isolate a subset of sites for
   which there is cursory evidence of differential methylation. Currently, the
   available screening approach adapts core routines from the
   [`limma`](http://bioconductor.org/packages/limma) R package. Following the
   style of the function for performing screening via `limma`, users may write
   their own screening functions and are invited to contribute such functions to
   the core software package by opening pull requests at the GitHub repository:
   \url{https://github.com/nhejazi/methyvim}.

2. Nonparametric estimates of VIMs, for the specified target parameter, are
   computed at each of the CpG sites passing the screening step. The VIMs are
   defined in such a way that the estimated effects is of an binary treatment on
   the methylation status of a target CpG site, controlling for the observed
   methylation status of the neighbors of that site. Currently, routines are
   adapted from the [`tmle`](https://CRAN.R-project.org/package=tmle) R package.

3. Since pre-screening is performed prior to estimating VIMs, we apply the
   modified marginal Benjamini and Hochberg step-up False Discovery Rate
   controlling procedure for multi-stage analyses (FDR-MSA), which is
   well-suited for avoiding false positive discoveries when testing is only
   performed on a subset of potential targets.

@@

#### Future Work

Support for continuous-valued treatments or exposures is _planned and ongoing_
but not yet available in stable releases of the `methyvim` software package.
Future releases will allow users to assess continuous-valued treatments by
relying on parameters estimable through implementations available in the
following software packages:

@@

The `show` method of the `methytmle` class summarizes a selection of the above
information for the user while masking some of the wealth of information given
when calling the same method for `GenomicRatioSet`. All information contained in
`GenomicRatioSet` objects is preserved in `methytmle` objects, so as to easen
interoperability with other differential methylation software for experienced
users. We refer the reader to the package vignette, "`methyvim`: Targeted
Data-Adaptive Estimation and Inference for Differential Methylation Analysis,"
included in any distribution of the software package, for further details.

@@

```{r setup-minfidata, message=FALSE}
suppressMessages(
  # numerous messages displayed at time of loading
  library(minfiData)
)
data(MsetEx)
mset <- mapToGenome(MsetEx)
grs <- ratioConvert(mset)
grs
```
We can create an object of class `methytmle` from any `GenomicRatioSet` object
simply invoking the S4 class constructor `.methytmle`:

@@

```{r fdr-msa}
fdr_p <- fdr_msa(pvals = vim(methyvim_cancer_ate)$pval,
                 total_obs = nrow(methyvim_cancer_ate))
```

@@

```{r methyvim-pvals-raw, fig.height=4, fig.width=6}
plot(methyvim_cancer_ate, type = "raw_pvals")
```

```{r methyvim-pvals-fdr, fig.height=4, fig.width=6}
plot(methyvim_cancer_ate, type = "fdr_pvals")
```

@@

```{r methyvim-volcano, fig.height=4, fig.width=6}
methyvolc(methyvim_cancer_ate)
```

@@

```{r methyvim-heatmap, fig.height=4, fig.width=6}
methyheat(methyvim_cancer_ate, smooth.heat = TRUE, left.label = "none")
```

@@

No competing interests were disclosed at the time of publication.
